---
title: "Homework 6"
author: "Sarah Younes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
set.seed(1)
```

## Problem 1

First, I will import the data.

```{r raw data}
homicide_data =
```

Next, I will import and clean the data, including filtering the data to necessary items.

```{r cleaning data}
homicide_data =
  read.csv(
    "./data/homicide-data.csv",
    na = "Unknown") |>
  janitor::clean_names() |>
  drop_na() |>
  mutate(
    victim_last = str_to_title(victim_last),
    victim_first = str_to_title(victim_first),
    city_state = paste(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    status = case_when(
      disposition == "Closed without arrest" ~ "Solved",
      disposition == "Closed by arrest" ~ "Solved",
      disposition == "Open/No arrest" ~ "Unsolved")) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, OK"))) |>
  filter(victim_race == "White" | victim_race == "Black")
```

Now, I will fit a logistic regression for Baltimore, MD with `status` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. I will tidy the output and obtain an estimate of the odds ratio.

```{r}
status_fit =
  glm(status ~ victim_age + victim_sex + victim_race, data = homicide_data, family = "binomial") |>
  broom::tidy() |>
  mutate(OR = exp(estimate))

status_fit
```

## Problem 2

First, I will download and clean the weather data set as instructed.

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Now, I will create a simple linear regression with `tmax` as the outcome and `tmin` and `prcp` as the predictors.

```{r}
tmax_fit = lm(tmax ~ tmin + prcp, data = weather_df) |>
  broom::tidy()

tmax_fit
```

## Problem 3