---
title: "Homework 6"
author: "Sarah Younes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(tidyverse)
set.seed(1)
```

## Problem 1

First, I will import and clean the data, including filtering the data to necessary items.

```{r cleaning homicide data}
homicide_data =
  read.csv(
    "./data/homicide-data.csv",
    na = "Unknown") |>
  janitor::clean_names() |>
  drop_na() |>
  mutate(
    victim_last = str_to_title(victim_last),
    victim_first = str_to_title(victim_first),
    city_state = paste(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    status = case_when(
      disposition == "Closed without arrest" ~ "Solved",
      disposition == "Closed by arrest" ~ "Solved",
      disposition == "Open/No arrest" ~ "Unsolved")) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, OK"))) |>
  filter(victim_race == "White" | victim_race == "Black")
```

Now, I will fit a logistic regression for Baltimore, MD with `status` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. I will tidy the output and obtain an estimate of the odds ratio.

```{r logistic regression}
regression_data =
  homicide_data |>
  mutate(
    status = case_when(
      status == "Solved" ~ 1,
      status == "Unsolved" ~ 0))

status_fit =
  glm(status ~ victim_age + victim_sex + victim_race, data = regression_data, family = "binomial") |>
  broom::tidy() |>
  mutate(OR = exp(estimate))

status_fit
```

## Problem 2

First, I will download the weather data set as instructed.

```{r downloading weather data, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Now, I will create a simple linear regression with `tmax` as the outcome and `tmin` and `prcp` as the predictors.

```{r linear regression}
tmax_fit = lm(tmax ~ tmin + prcp, data = weather_df)
```

Now, I will pull r-squared.

```{r}
r2 =
  tmax_fit |>
  broom::glance() |>
  select(r.squared)

r2
```

Now, I will pull β̂ 1 and β̂ 2.

```{r}
betas =
  tmax_fit |>
  broom::tidy() |>
  select(term, estimate)

beta1 = 
  betas |>
  filter(term == "tmin") |>
  select(estimate)

beta2 =
  betas |>
  filter(term == "prcp") |>
  select(estimate)

log_beta1_beta2 = log(abs(beta1 * beta2))
```

I will draw 5,000 bootstrap samples and produce estimates of log(β̂ 1∗β̂ 2) using `broom::tidy()` and `pivot_wider`.

```{r}
prcp_tmin_bootstrap =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(weather_df) lm(tmax ~ tmin + prcp, data = weather_df)),
    results = map(models, broom::tidy),
    sample_number = row_number()) |>
  unnest(results) |>
  filter(term %in% c("prcp", "tmin")) |>
  pivot_wider(
    id_cols = sample_number,
    names_from = term,
    values_from = estimate) |>
  mutate(
    log = log(abs(tmin * prcp)))

prcp_tmin_bootstrap
```

I will draw 5,000 bootstrap samples and produce estimates of r-squared using `broom::glance`.

```{r}
r2_bootstrap =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(weather_df) lm(tmax ~ tmin + prcp, data = weather_df)),
    results = map(models, broom::glance),
    sample_number = row_number()) |>
  unnest(results) |>
  select(sample_number, r.squared)

r2_bootstrap
```

Now, I will identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r-squared and the log of B1 and B2.

## Problem 3

I will import the birthweight data set.

```{r}
birthweight_df =
  read_csv("./data/birthweight.csv")
```

I will use a data-driven modeling process to propose a regression model for birthweight. I compare my findings from  two methods: first, I will create ggplots to look at trends in each variable against `bwt` on the y-axis and look at the strongest visual trends, and second, I will run statistical tests (correlations, t-tests, or ANOVAs depending on the type of independent variable) against `bwt` as the dependent variable and look at the strongest numeric correlations or largest test statistic. I expect to see consistency between the two methods (ggplots and statistics) and will include the variables with the strongest visual or numeric relationships in my model.

However, I am automatically excluding `pnumlbw` and `pnumsga` as potential independent variables because their values across the entire sample were all 0, so there cannot be any trends between them and `bwt`.

Here are my ggplots:

```{r ggplots}
birthweight_df |>
  ggplot(aes(x = babysex, y = bwt, group = babysex)) + geom_boxplot()

birthweight_df |>
  ggplot(aes(x = bhead, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = blength, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = delwt, y = bwt)) + geom_point() + geom_smooth(method = "lm")

birthweight_df |>
  ggplot(aes(x = fincome, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = frace, y = bwt, group = frace)) + geom_violin()

birthweight_df |>
  ggplot(aes(x = gaweeks, y = bwt)) + geom_point() + geom_smooth(method = "lm")

birthweight_df |>
  ggplot(aes(x = malform, y = bwt, group = malform)) + geom_boxplot()

birthweight_df |>
  ggplot(aes(x = menarche, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = mheight, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = momage, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = mrace, y = bwt, group = mrace)) + geom_violin()

birthweight_df |>
  ggplot(aes(x = parity, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = ppbmi, y = bwt)) + geom_point() + geom_smooth(method = "lm")

birthweight_df |>
  ggplot(aes(x = ppwt, y = bwt)) + geom_point() + geom_smooth(method = "lm")

birthweight_df |>
  ggplot(aes(x = smoken, y = bwt)) + geom_point()

birthweight_df |>
  ggplot(aes(x = wtgain, y = bwt)) + geom_point() + geom_smooth(method = "lm")
```

Here are my statistical tests:

```{r correlations}
cor.test(birthweight_df$bwt, birthweight_df$bhead, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$blength, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$delwt, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$fincome, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$gaweeks, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$menarche, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$mheight, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$momage, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$parity, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$ppbmi, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$ppwt, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$smoken, method = "pearson") |> broom::tidy()

cor.test(birthweight_df$bwt, birthweight_df$wtgain, method = "pearson") |> broom::tidy()

t.test(birthweight_df$bwt, birthweight_df$babysex) |> broom::tidy()

t.test(birthweight_df$bwt, birthweight_df$malform) |> broom::tidy()

aov(birthweight_df$bwt ~ birthweight_df$frace) |> broom::tidy()

aov(birthweight_df$bwt ~ birthweight_df$mrace) |> broom::tidy()
```

The following variables showed both high visual relationships from ggplots and high statistical relationships (correlations above r-squared = 0.7 or p-values below alpha = 0.05): bhead, blength, babysex, malform, frace, and mrace.

```{r}
model = lm(bwt ~ bhead + blength, data = birthweight_df)
```

*Add-predictions & add_residuals

```{r}
birthweight_df |>
  modelr::add_residuals(model) |>
  ggplot(aes(x = bwt, y = resid)) + geom_point() + geom_smooth()
```

Now, I will compare my model to two others:

```{r}
model_1 = lm(bwt ~ blength + gaweeks, data = birthweight_df)

model_2 = lm(bwt ~ bhead + blength + babysex + (bhead * blength * babysex), data = birthweight_df)
```